

# Results and Discussion

`r if(gitbook_on){knitr::asis_output("Results of Phase 1 and Phase 2 assessments are summarized in Figure \\@ref(fig:map-interactive) with additional details provided in sections below.")}`

<br>

```{r map-interactive, fig.cap='Map of fish passage and habitat confirmation results', eval=gitbook_on}


##make colors for the priorities
pal <- 
   leaflet::colorFactor(palette = c("red", "yellow", "grey", "black"), 
               levels = c("High", "Moderate", "Low", "No Fix"))

pal_phase1 <- 
   leaflet::colorFactor(palette = c("red", "yellow", "grey", "black"), 
               levels = c("High", "Moderate", "Low", NA))


  
map <- leaflet::leaflet(height=500, width=780) |>
  leaflet::addTiles()  |> 
  # leafem::addMouseCoordinates(proj4 = 26911) |> ##can't seem to get it to render utms yet
  # leaflet::addProviderTiles(providers$"Esri.DeLorme") |> 
  leaflet::addProviderTiles("Esri.WorldTopoMap", group = "Topo") |>
  leaflet::addProviderTiles("Esri.WorldImagery", group = "ESRI Aerial") |> 
  # leaflet::addPolygons(data = wshd_study_areas, color = "#F29A6E", weight = 1, smoothFactor = 0.5,
  #             opacity = 1.0, fillOpacity = 0,
  #             fillColor = "#F29A6E", label = wshd_study_areas$watershed_group_name) |>
  leaflet::addPolygons(data = wshds, color = "#0859C6", weight = 1, smoothFactor = 0.5,
              opacity = 1.0, fillOpacity = 0.25,
              fillColor = "#00DBFF",
              label = wshds$stream_crossing_id,
              popup = leafpop::popupTable(x = dplyr::select(wshds |> sf::st_set_geometry(NULL),
                                                     Site = stream_crossing_id,
                                                     elev_site:area_km),
                                          feature.id = F,
                                          row.numbers = F),
              group = "Phase 2") |> 
  leaflet::addLegend(
    position = "topright",
    colors = c("red", "yellow", "grey", "black"),
    labels = c("High", "Moderate", "Low", 'No fix'), opacity = 1,
    title = "Fish Passage Priorities") |> 
  
leaflet::addCircleMarkers(data=dplyr::filter(tab_map_phase_1, stringr::str_detect(source, 'phase1') | stringr::str_detect(source, 'pscis_reassessments')),
                 label = dplyr::filter(tab_map_phase_1, stringr::str_detect(source, 'phase1') | stringr::str_detect(source, 'pscis_reassessments')) |> dplyr::pull(pscis_crossing_id),
                 # label = tab_map_phase_1$pscis_crossing_id,
                 labelOptions = leaflet::labelOptions(noHide = F, textOnly = TRUE),
                 popup = leafpop::popupTable(x = dplyr::select((tab_map_phase_1 |> sf::st_set_geometry(NULL) |> dplyr::filter(stringr::str_detect(source, 'phase1') | stringr::str_detect(source, 'pscis_reassessments'))),
                                                        Site = pscis_crossing_id, Priority = priority_phase1, Stream = stream_name, Road = road_name, `Habitat value`= habitat_value, `Barrier Result` = barrier_result, `Culvert data` = data_link, `Culvert photos` = photo_link, `Model data` = model_link),
                                             feature.id = F,
                                             row.numbers = F),
                 radius = 9,
                 fillColor = ~pal_phase1(priority_phase1),
                 color= "#ffffff",
                 stroke = TRUE,
                 fillOpacity = 1.0,
                 weight = 2,
                 opacity = 1.0,
                 group = "Phase 1") |> 
  # leaflet::addPolylines(data=habitat_confirmation_tracks,
  #              opacity=0.75, color = '#e216c4',
  #              fillOpacity = 0.75, weight=5, group = "Phase 2") |>

leaflet::addAwesomeMarkers(
    lng = as.numeric(photo_metadata$gps_longitude),
    lat = as.numeric(photo_metadata$gps_latitude),
    popup = leafpop::popupImage(photo_metadata$url, src = "remote"),
    clusterOptions = leaflet::markerClusterOptions(),
    group = "Phase 2") |>


  leaflet::addCircleMarkers(
    data=tab_map_phase_2,
    label = tab_map_phase_2$pscis_crossing_id,
    labelOptions = leaflet::labelOptions(noHide = T, textOnly = TRUE),
    popup = leafpop::popupTable(x = dplyr::select((tab_map_phase_2 |> sf::st_drop_geometry()),
                                           Site = pscis_crossing_id,
                                           Priority = priority,
                                           Stream = stream_name,
                                           Road = road_name,
                                           `Habitat (m)`= upstream_habitat_length_m,
                                           Comments = comments,
                                           `Culvert data` = data_link,
                                           `Culvert photos` = photo_link,
                                           `Model data` = model_link),
                                feature.id = F,
                                row.numbers = F),
    radius = 9,
    fillColor = ~pal(priority),
    color= "#ffffff",
    stroke = TRUE,
    fillOpacity = 1.0,
    weight = 2,
    opacity = 1.0,
    group = "Phase 2"
    ) |>

  leaflet::addLayersControl(
    baseGroups = c(
      "Esri.DeLorme",
      "ESRI Aerial"),
    overlayGroups = c("Phase 1", "Phase 2"),
    options = leaflet::layersControlOptions(collapsed = F)) |>
  leaflet.extras::addFullscreenControl() |>
  leaflet::addMiniMap(tiles = leaflet::providers$"Esri.NatGeoWorldMap",
             zoomLevelOffset = -6, width = 100, height = 100)

map

```


## Site Assessment Data Since 2020

Fish passage assessment procedures conducted through SERNbc in the Upper Fraser River Watershed since 2023 are amalgamated  
`r if(identical(gitbook_on, FALSE)){knitr::asis_output(paste0("online within the Results and Discussion section of the report found ", ngr::ngr_str_link_url(url_base = params$report_url, anchor_text = "here")))}else{knitr::asis_output("in Tables \\@ref(tab:tab-sites-sum-cap) - \\@ref(tab:tab-sites-cap).")}`

<br>

```{r ld-db}

conn = fpr::fpr_db_conn()

sites_all <- fpr::fpr_db_query(
  query = "SELECT * FROM working.fp_sites_tracking"
)

DBI::dbDisconnect(conn)

```

```{r summarize-sites}

# unique(sites_all$watershed_group_name)
# 
wsg <- c(
  "Lower Chilako River",
  "Willow River",
  "Tabor River",
  "Upper Fraser River",
  "Nechako River",
  "Morkill River",
  "Francois Lake"
)

# more straight forward is new graph only watersheds
# wsg_ng <- "Elk River"

# here is a summary with Elk watershed group removed
sites_all_summary <- sites_all |> 
  # make a flag column for uav flights
  dplyr::mutate(
    uav = dplyr::case_when(
      !is.na(link_uav1) ~ "yes",
      T ~ NA_character_
    )) |> 
  # remove the elk counts
  dplyr::filter(watershed_group %in% wsg) |>
  dplyr::group_by(watershed_group) |> 
  dplyr::summarise(
    dplyr::across(assessment:fish_sampling, ~ sum(!is.na(.x))),
    uav = sum(!is.na(uav))
  ) |> 
  sf::st_drop_geometry() |> 
  # make pretty names
  dplyr::rename_with(~ stringr::str_replace_all(., "_", " ") |> 
                       stringr::str_to_title()) |> 
  # annoying special case
  dplyr::rename(
    `Drone Imagery` = Uav) |> 
  janitor::adorn_totals()

```

```{r tab-sites-sum-cap, results="asis"}
my_caption = "Summary of fish passage assessment procedures conducted in the Upper Fraser River watershed through SERNbc since 2023."
my_tab_caption(tip_flag = FALSE)
```

```{r tab-sites-sum}
sites_all_summary |> 
  dplyr::mutate(across(everything(), as.character)) |> 
  my_dt_table(
    page_length = 20,
    cols_freeze_left = 0
  )
```

<br>


```{r tab-sites-cap, results="asis"}
my_caption = "Details of fish passage assessment procedures conducted in the Upper Fraser River watershed through SERNbc since 2023."
my_tab_caption()
```

```{r tab-sites-all}
sites_all |>
  sf::st_drop_geometry() |>
  dplyr::filter(watershed_group %in% wsg) |>
  dplyr::relocate(watershed_group, .after = my_crossing_reference) |> 
  dplyr::select(-idx) |>
  # make pretty names
  dplyr::rename_with(~ . |>
                       stringr::str_replace_all("_", " ") |>
                       stringr::str_replace_all("repo", "Report") |>
                       stringr::str_replace_all("uav", "Drone") |>
                       stringr::str_to_title()) |> 
  # dplyr::arrange(desc(stream_crossing_id)) |> 
  # make all the columns strings so we can filter them 
  dplyr::mutate(across(everything(), as.character)) |> 
  
  my_dt_table(
    cols_freeze_left = 1,
    escape = FALSE
  )
```

<br>



## Collaborative GIS Environment

In addition to numerous layers documenting fieldwork activities since 2023, a summary of background information spatial layers and tables loaded to the collaborative GIS project (`r params$gis_project_name`) at the
time of writing (`r format(lubridate::now(), "%Y-%m-%d")`) are included 
`r if(identical(gitbook_on, FALSE)){knitr::asis_output(paste0("online ", ngr::ngr_str_link_url(url_base = params$report_url, anchor_text = "here")))}else{knitr::asis_output("in Table \\@ref(tab:tab-rfp-tracking).")}`.

<br>

```{r rfp-metadata, eval = params$update_gis}
# grab the metadata
md <- rfp::rfp_meta_bcd_xref()

# burn locally so we don't nee to wait for it
md |> 
  readr::write_csv("data/rfp_metadata.csv")
```


```{r rfp-tracking-copy, eval = params$update_gis}

md_raw <- readr::read_csv("data/rfp_metadata.csv")

md <- dplyr::bind_rows(
  md_raw,
  rfp::rfp_xref_layers_custom
) 
                    
# first we will copy the doc from the Q project to this repo - the location of the Q project is outside of the repo!!
q_path_stub <- fs::path_expand(fs::path("~/Projects/gis", params$gis_project_name))

# this is differnet than Neexdzii Kwa as it lists layers vs tracking file (tracking file is newer than this project).  
# could revert really easily to the tracking file if we wanted to.
gis_layers_ls <- sf::st_layers(fs::path(q_path_stub, "background_layers.gpkg")) 

gis_layers <- tibble::tibble(content = gis_layers_ls[["name"]])
   
# remove the `_vw` from the end of content
rfp_tracking_prep <- dplyr::left_join(
  gis_layers |> 
    dplyr::distinct(content, .keep_all = FALSE),
  
  md |> 
    dplyr::select(content = object_name, url = url_browser, description),
  
  by = "content"
) |> 
  dplyr::arrange(content)

rfp_tracking_prep |> 
  readr::write_csv("data/rfp_tracking_prep.csv")

```

```{r tab-rfp-tracking, eval= gitbook_on}
rfp_tracking_prep <- readr::read_csv(
  "data/rfp_tracking_prep.csv"
) 

rfp_tracking_prep |> 
  fpr::fpr_kable(caption_text = "Layers loaded to collaborative GIS project.",
                 footnote_text = "Metadata information for bcfishpass and bcfishobs layers can be provided here in the future but  currently can usually be sourced from https://smnorris.github.io/bcfishpass/06_data_dictionary.html .",
                 scroll = gitbook_on)



```


## Phase 1 - Fish Passage Assessemnts

During the 2023 field season, field assessments were conducted between `r format(min(pscis_phase1_2023$date), format="%B %d, %Y")` and `r format(max(pscis_phase1_2023$date), format="%B %d, %Y")` by Allan Irvine, R.P.Bio. and Mateo Winterscheidt, B.Sc. A total of `r pscis_phase1_2023 |> nrow()` Phase 1 Fish Passage Assessments were completed.

<br>

During the 2024 field season, field assessments were conducted between `r format(min(pscis_phase1_2024$date), format="%B %d, %Y")` and `r format(max(pscis_phase1_2024$date), format="%B %d, %Y")` by Allan Irvine, R.P.Bio. and Lucy Schick, B.Sc. A total of 9 Fish Passage Assessments were completed, including `r n_distinct(pscis_phase1_2024$my_crossing_reference)` Phase 1 assessments and `r pscis_reassessments |> nrow()` reassessments.

<br>

Of the `r pscis_all |> filter(source != 'pscis_phase2.xlsm') |> nrow()` sites where fish passage assessments were completed, `r pscis_phase1 |> nrow()` were not yet inventoried in the PSCIS system.  This included `r pscis_phase1 |> filter(barrier_result == 'Passable') |> nrow()` crossings considered "passable", `r pscis_phase1 |> filter(barrier_result == 'Potential') |> nrow() |> english::as.english()` crossings considered "potential" barriers, and `r pscis_phase1 |> filter(barrier_result == 'Barrier') |> nrow()` crossings considered "barriers" according to threshold values based on culvert embedment, outlet drop, slope, diameter (relative to channel size) and length [@moe2011Fieldassessment]. Additionally, although all were considered fully passable, `r pscis_phase1 |> filter(barrier_result == 'Unknown') |> nrow() |> english::as.english()` crossings assessed were fords and were ranked as "unknown" according to the provincial protocol.

<br>

Reassessments were completed at `r pscis_reassessments |> nrow()` sites where PSICS data required updating.

<br>

A summary of crossings assessed, a rough cost estimate for remediation, and a priority ranking for follow-up for Phase 1 sites is presented in Table \@ref(tab:cost-est-phase-1). Detailed data with photos are presented in `r if(gitbook_on){knitr::asis_output("[Appendix - Phase 1 Fish Passage Assessment Data and Photos]")} else knitr::asis_output(ngr::ngr_str_link_url(url_base = params$report_url, url_resource = "appendix---phase-1-fish-passage-assessment-data-and-photos.html", anchor_text = "Attachment 1"))`.

<br>

The "Barrier" and "Potential Barrier" rankings used in this project followed @moe2011Fieldassessment and represent an assessment of passability for juvenile salmon or small resident rainbow trout under any flow conditions that may occur throughout the year [@clarkinNationalInventoryAssessment2005; @bellFisheriesHandbookEngineering1991; @thompsonAssessingFishPassage2013]. As noted in @bourne_etal2011Barriersfish, with a detailed review of different criteria in @kemp_ohanley2010Proceduresevaluating, passability of barriers can be quantified in many different ways. Fish physiology (i.e. species, length, swim speeds) can make defining passability complex but with important implications for evaluating connectivity and prioritizing remediation candidates [@bourne_etal2011Barriersfish; @shaw_etal2016Importancepartial; @mahlum_etal2014EvaluatingBarrier; @kemp_ohanley2010Proceduresevaluating].  @washingtondepartmentoffishwildlife2009FishPassage present criteria for assigning passability scores to culverts that have already been assessed as barriers in coarser level assessments. These passability scores provide additional information to feed into decision making processes related to the prioritization of remediation site candidates and have potential for application in British Columbia.  


`r if(gitbook_on){knitr::asis_output("")} else knitr::asis_output("<br><br><br><br>")`

<br>

<!-- turn on once we have pscis data -->

```{r cost-est-phase-1, eval = TRUE}
 tab_cost_est_phase1 |>
  select(`PSCIS ID`:`Cost Est ( $K)`) |> 
   fpr::fpr_kable(caption_text = paste0("Upstream habitat estimates and cost benefit analysis for Phase 1 assessments ranked as a 'barrier' or 'potential' barrier. ", sp_network_caption),
                  scroll = gitbook_on)
 
```

<br>

## Phase 2 - Habitat Confirmation Assessments

During the 2024 field assessments, habitat confirmation assessments were conducted at `r habitat_confirmations_priorities |> dplyr::distinct(stream_name) |> nrow()` sites within the Nechako River, Lower Chilako River, Upper Fraser River, Morkill River, and Francois Lake watershed groups. A total of approximately `r habitat_confirmations_priorities |> pull(length_surveyed) |> sum(na.rm = TRUE) |> round(-3)/1000` km of stream was assessed. Georeferenced field maps are provided in `r knitr::asis_output(if (gitbook_on) "[Attachment 1](#attach-1-maps)" else ngr::ngr_str_link_url(url_base = params$report_url, url_resource = "attachment-1---maps.html", anchor_text = "Attachment 1"))`.

<br>

As collaborative decision making was ongoing at the time of reporting, site prioritization can be considered preliminary. 
Results are summarized in `r if(gitbook_on){knitr::asis_output("Figure \\@ref(fig:map-interactive) and ")}` Table \@ref(tab:tab-overview) with raw habitat data included in `r knitr::asis_output(if (gitbook_on) "[Attachment 2](#raw-dat)" else ngr::ngr_str_link_url(url_base = params$report_url, url_resource = "attachment-2---habitat-assessment-and-fish-sampling-data.html", anchor_text = "Attachment 2"))`. A summary of preliminary modeling results illustrates the estimated Chinook, coho, and steelhead spawning and rearing habitat potentially available upstream of each crossing, based on measured/modelled channel width and upstream accessible stream length, as presented in Figure \@ref(fig:plot-model-all). Detailed information for each site assessed with Phase 2 assessments (including maps) are presented within site specific appendices to this document.

`r if(gitbook_on){knitr::asis_output("")} else knitr::asis_output("<br><br><br><br>")`

<br>


```{r tab-overview, eval = TRUE}
table_phase2_overview <- function(dat, caption_text = '', font = font_set, scroll = TRUE){
  dat2 <- dat |>
    kable(caption = caption_text, booktabs = T, label = NA) |>
    kableExtra::kable_styling(c("condensed"),
                              full_width = T,
                              font_size = font) |>
    kableExtra::column_spec(column = c(10), width_min = '1.5in') |>
    kableExtra::column_spec(column = c(1:9), width_max = '1in')

  if(identical(scroll,TRUE)){
    dat2 <- dat2 |>
      kableExtra::scroll_box(width = "100%", height = "500px")
  }
  dat2
}

tab_overview |> 
  table_phase2_overview(caption_text = paste0("Overview of habitat confirmation sites. ", sp_rearing_caption), 
                        scroll = gitbook_on)

```

`r if(gitbook_on){knitr::asis_output("<br>")} else knitr::asis_output("\\pagebreak")`


```{r phase-2-cv-assess-sum}
fpr::fpr_table_cv_summary(dat = pscis_phase2)  |>  
  fpr::fpr_kable(caption_text = 'Summary of Phase 2 fish passage reassessments.', scroll = gitbook_on)
```


`r if(gitbook_on){knitr::asis_output("<br>")} else knitr::asis_output("\\pagebreak")`


```{r cost-est-phase-2, eval=T}
tab_cost_est_phase2 |>
    dplyr::rename(
    `PSCIS ID` = pscis_crossing_id,
    Stream = stream_name,
    Road = road_name,
    `Barrier Result` = barrier_result,
    `Habitat value` = habitat_value,
    `Habitat Upstream (m)` = upstream_habitat_length_m,
    `Stream Width (m)` = avg_channel_width_m,
    Fix = crossing_fix_code,
    `Cost Est ( $K)` = cost_est_1000s,
    `Cost Benefit (m / $K)` = cost_net,
    `Cost Benefit (m2 / $K)` = cost_area_net
  ) |> 
  fpr::fpr_kable(caption_text = paste0("Cost benefit analysis for Phase 2 assessments. ", sp_rearing_caption),
                 scroll = gitbook_on)

```


`r if(gitbook_on){knitr::asis_output("<br>")} else knitr::asis_output("\\pagebreak")`

```{r tab-habitat-summary}
tab_hab_summary |> 
  dplyr::filter(stringr::str_like(Location, 'upstream')) |> 
  dplyr::select(-Location) |> 
  dplyr::rename(`PSCIS ID` = Site, `Length surveyed upstream (m)` = `Length Surveyed (m)`) |>
  fpr::fpr_kable(caption_text = 'Summary of Phase 2 habitat confirmation details.', scroll = gitbook_on)


```

<br>

```{r tab-wshd-sum-all}
fpr::fpr_table_wshd_sum() |> 
  fpr::fpr_kable(caption_text = paste0('Summary of watershed area statistics upstream of Phase 2 crossings.'),
                                 footnote_text = 'Elev P60 = Elevation at which 60% of the watershed area is above', scroll = gitbook_on)

```

<br>

```{r plot-model-all-cap}

my_caption = 'Summary of potential rearing and spawning habitat upstream of habitat confirmation assessment sites. See Table \\@ref(tab:tab-fish-spawning-rearing) for modelling thresholds.'

```


```{r plot-model-all, fig.cap= my_caption, eval=T}


bcfp_xref_plot <- xref_bcfishpass_names |> 
  dplyr::filter(!is.na(id_join) &
         !stringr::str_detect(bcfishpass, 'below') &
         !stringr::str_detect(bcfishpass, 'all') &
         !stringr::str_detect(bcfishpass, '_ha') &
         (stringr::str_detect(bcfishpass, 'rearing') | 
          stringr::str_detect(bcfishpass, 'spawning')))


bcfishpass_phase2_plot_prep <- bcfishpass |> 
  dplyr::mutate(dplyr::across(where(is.numeric), round, 1)) |>
  dplyr::filter(stream_crossing_id %in% (pscis_phase2 |> dplyr::pull(pscis_crossing_id))) |> 
  dplyr::select(stream_crossing_id, dplyr::all_of(bcfp_xref_plot$bcfishpass)) |> 
  dplyr::mutate(stream_crossing_id = as.factor(stream_crossing_id)) |> 
  tidyr::pivot_longer(cols = ch_spawning_km:st_rearing_km) |> 
  dplyr::filter(value > 0.0 &
                !is.na(value)) |> 
  dplyr::mutate(
    # name = dplyr::case_when(stringr::str_detect(name, '_rearing') ~ paste0(model_species_name, " rearing km"),
    #                         TRUE ~ name),
    # name = dplyr::case_when(stringr::str_detect(name, '_spawning') ~ paste0(model_species_name, " spawning km"),
    #                         TRUE ~ name),
    # Use when more than one modelling species 
   name = stringr::str_replace_all(name, '_rearing', ' rearing'),
   name = stringr::str_replace_all(name, '_spawning', ' spawning')
  )

    
  
bcfishpass_phase2_plot_prep |> 
  ggplot2::ggplot(ggplot2::aes(x = stream_crossing_id, y = value)) +
  ggplot2::geom_bar(stat = "identity") +
  ggplot2::facet_wrap(~name, ncol = 2) +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  ggplot2::labs(x = "Site", y = "Modelled habitat (km)")+
  ggplot2::theme_light(base_size = 11)


```
