---
title: "photos import"
date: "Created: 2024-05-17 | Updated: `r format(Sys.Date(), '%Y-%m-%d')`"
output: 
  html_document:
    code_folding: "hide"
params:
  repo_owner: "NewGraphEnvironment"
  repo_name: "fish_passage_fraser_2023_reporting"
  gis_name: "sern_fraser_2024"
  job_name: "2024-074-sern-fraser-fish-passage"
---

```{r setup, eval= TRUE}
knitr::opts_chunk$set(echo=TRUE, message=FALSE, warning=FALSE, dpi=60, out.width = "100%", eval = FALSE)
options(scipen=999)
options(knitr.kable.NA = '--') #'--'
options(knitr.kable.NAN = '--')

```

```{r paths, eval=TRUE}

# 2024 data photos
dir_server_photos <- fs::path_expand(fs::path("~/Library/CloudStorage/OneDrive-Personal/Projects", params$job_name, "data/photos/sorted"))

dir_repo_photos <- fs::path_expand(fs::path("~/Projects/repo/", params$repo_name, "/data/photos/"))
```

This is a simple script to bulk import all photos tagged with `_k_` or `crossings_all`  from a shared server to the local repo directory. Idea here is that all collaborators on raw project (who have access to the shared server) can add photos to the shared server and alter
photos on the shared server (ie. organize, rename) and then update the repo. 

<br>

Naming conventions to allow `git` to track photos are included in the project `.gitignore` (ex. `_k_` included or photo
is named `crossing_all.JPG`).

<br>

This script keeps things very simple by first deleting all the photos locally, and then copying only the photos tagged with `_k_` or `crossings_all` from the shared server to the local repo. Fraser 2024 - no deleting because we also need to add photos from 2023 which makes things funky - see https://github.com/NewGraphEnvironment/fish_passage_fraser_2023_reporting/issues/129

```{r copy, eval=TRUE}

# create the new mother photos directory
fs::dir_create(dir_repo_photos)

# These are the photos we are copying
photos_to_copy <- fs::dir_ls(dir_server_photos, recurse = TRUE, type = "file") %>% 
  .[stringr::str_detect(., "_k_|crossing_all")]

# Make the paths for where we are copying the photos to
files_to_paste <- stringr::str_remove(photos_to_copy, paste0("^", dir_server_photos, "/"))
photos_to_paste <- fs::path(dir_repo_photos, files_to_paste)

# Create directories for the new file paths
dirs_to_create <- files_to_paste |>
  fs::path_dir() |>
  unique()

fs::dir_create(fs::path(dir_repo_photos, "/", dirs_to_create))


# Copy photos from the shared server to the local repo
purrr::map2(photos_to_copy, photos_to_paste, ~ fs::file_copy(.x, .y, overwrite = TRUE))
```

```{r 2023-photos}

# This is unique to Fraser because we data and photos from 2023 and 2024 - see issue here https://github.com/NewGraphEnvironment/fish_passage_fraser_2023_reporting/issues/129

# So we also need to move over the photos from 2023 which are stored in mergin projects

# simpcw_2023 photo 
dir_photos_simpcw_2023 <- fs::path_expand("~/Projects/gis/sern_simpcw_2023/ignore_mobile/photos")

# These are the photos we are copying
photos_to_copy <- fs::dir_ls(dir_photos_simpcw_2023, recurse = TRUE, type = "file") %>% 
  .[stringr::str_detect(., "_k_|crossing_all")]

# Make the paths for where we are copying the photos to
files_to_paste <- stringr::str_remove(photos_to_copy, paste0("^", dir_photos_simpcw_2023, "/"))
photos_to_paste <- fs::path(dir_repo_photos, files_to_paste)

# Create directories for the new file paths
dirs_to_create <- files_to_paste |>
  fs::path_dir() |>
  unique()

fs::dir_create(fs::path(dir_repo_photos, "/", dirs_to_create))


# Copy photos from the shared server to the local repo
purrr::map2(photos_to_copy, photos_to_paste, ~ fs::file_copy(.x, .y, overwrite = TRUE))


```

