---
title: "0060-pscis-tidy"
date: "Created: 2024-06-20 | Updated: `r format(Sys.Date(), '%Y-%m-%d')`"
output: 
  html_document:
    code_folding: "hide"
params:
  repo_owner: "NewGraphEnvironment"
  repo_name: "fish_passage_skeena_2023_reporting"
editor_options: 
  chunk_output_type: console
---

```{r setup, echo=TRUE, include = TRUE}
knitr::opts_chunk$set(echo=TRUE, message=FALSE, warning=FALSE, dpi=60, out.width = "100%", eval = FALSE)
options(scipen=999)
options(knitr.kable.NA = '--') #'--'
options(knitr.kable.NAN = '--')
```

# Purpose of this script

This scripts is used to prepare data for QA in QGIS. It 

First we load the necessary packages.
```{r}
source('scripts/packages.R')
```



## Fetch and backup the data

Name the project directory we are pulling from and import the data. We use `fpr_sp_gpkg_backup` to import and backup the data, as well as update the UTMS. We will also write the data to a csv and Rdata file and commit for version control using git.

```{r import}
path <- "~/Projects/gis/sern_simpcw_2023"

form_pscis <- fpr_sp_gpkg_backup(
  path_gpkg = paste0(path, "/data_field/2023/form_pscis_2023.gpkg"),
  dir_backup = "data/backup/simpcw/",
  update_utm = TRUE,
  update_site_id = F,
  write_back_to_path = FALSE,
  write_to_csv = TRUE,
  write_to_rdata = TRUE,
  return_object = TRUE)

# pscic_2023 <- sf::st_read('/Users/lucyschick/Projects/gis/sern_simpcw_2023/data_field/2023/form_pscis_2023.gpkg') |> 
#   arrange(my_crossing_reference)
# 
# pscic_old <- sf::st_read('/Users/lucyschick/Projects/gis/sern_simpcw_2023/form_pscis.gpkg') |> 
#   arrange(my_crossing_reference)|> 
#   slice(2:(nrow(pscic_old) - 1)) |> 
#   mutate(my_crossing_reference = case_when(
#     my_crossing_reference == 1300094 ~ 13900094,
#     TRUE ~ my_crossing_reference
#   )) |> 
#   arrange(my_crossing_reference)
# 
# 
# setdiff(unique(pscic_2023$my_crossing_reference), unique(pscic_old$my_crossing_reference))
# 
# waldo::compare(pscic_2023, pscic_old, max_diffs = Inf)
  
```


## Clean and QA the data

First we check for duplicated sites. In this case, there are too sites with `my_crossing_reference` = 13905538, but they are not duplicates, just two crossings at the same location.

```{r dups}
form_pscis %>%
  filter(!is.na(my_crossing_reference)) %>%
  group_by(my_crossing_reference) %>%
  filter(n()>1)
```


We need to check for sites that have a culvert length (length_or_width_meters) over 99.9 or a fill depth (fill_depth_meters) over 9.9, anything over this will cause errors in submission sheet. If over the maximums, we will change them to 99.9 and 9.9, respectively, and will append a note to the assessment comments.


```{r max-length-fill}

# First lets check
form_pscis |> 
  filter(length_or_width_meters > 99.9 | fill_depth_meters > 9.9)

# Now we will change them to 99.9 and 9.9, respectively, and will append a note to the assessment comments.
form_pscis <- form_pscis %>%
  dplyr::mutate(
    assessment_comment = case_when(
      length_or_width_meters > 99.9 ~ paste0(assessment_comment, 'Culvert length ', length_or_width_meters, 'm but changed to 99.9m to meet submission requirements.'),
      TRUE ~ assessment_comment
    ),
    length_or_width_meters = case_when(
      length_or_width_meters > 99.9 ~ 99.9,
      TRUE ~ length_or_width_meters
    ),
    assessment_comment = case_when(
      fill_depth_meters > 9.9 ~ paste0(assessment_comment, 'Fill depth ', fill_depth_meters, 'm but changed to 9.9m to meet submission requirements.'),
      TRUE ~ assessment_comment
    ),
    fill_depth_meters = case_when(
      fill_depth_meters > 9.9 ~ 9.9,
      TRUE ~ fill_depth_meters
    )
  )

```

We need to fix the times because they are in UTC and we need them in PDT. This issue is documented here https://github.com/NewGraphEnvironment/fish_passage_template_reporting/issues/18

```{r}

form_pscis <- form_pscis |> 
# make a new column for the time as is with different name then mutate to PST
  # we don't need the new column but will leave here for now so we can visualize and confirm the time is correct
  mutate(date_time_start_raw = date_time_start,
         date_time_start = lubridate::force_tz(date_time_start_raw, tzone = "America/Vancouver"),
         date_time_start = lubridate::with_tz(date_time_start, tzone = "UTC")) |> 
  relocate(date_time_start_raw, .after = date_time_start)

## Double check the time is correct and now remove the date_time_start_raw column
form_pscis <- form_pscis |> 
  select(-date_time_start_raw)
```



Now we will:
- remove the site used to make the form
- split date time column into date and time
- Change all yes_no columns that have NA values to 'No', This should be done in the form as per https://github.com/NewGraphEnvironment/dff-2022/issues/119#issuecomment-1781242709
- back up the original assessment comments so we can redo this amalgamation of text if we need to
- Add additional columns and rearrange. In future we should have all of these columns already in the form as per https://github.com/NewGraphEnvironment/dff-2022/issues/119#issuecomment-1781242709

```{R }
form_pscis_cleaned <- form_pscis %>%
  # remove the site used to make the form
  filter(site_id != '12345' | !is.na(date_time_start)) %>%

  #split date time column into date and time
  dplyr::mutate(
    date_time_start = lubridate::ymd_hms(date_time_start),
    date = lubridate::date(date_time_start),
    time = hms::as_hms(date_time_start),

    # This should be done in the form as per https://github.com/NewGraphEnvironment/dff-2022/issues/119#issuecomment-1781242709
    across(contains('yes_no'), ~replace_na(.,'No')),

    # some numeric fields for CBS have NA values when a user input 0 https://github.com/NewGraphEnvironment/dff-2022/issues/119#issuecomment-1781242709
    # this could lead to errors as is. Should be fixed in form if possible
    across(c(outlet_drop_meters, outlet_pool_depth_0_01m, culvert_slope_percent, stream_slope),
           ~case_when(
             crossing_type == 'Closed Bottom Structure' ~
               replace_na(.,0),
             TRUE ~ .
           )),
    stream_name = str_replace_all(stream_name, 'Trib ', 'Tributary '),
    road_name = str_replace_all(road_name, 'Hwy ', 'Highway '),
    # back up the original assessment comments so we can redo this amalgamation of text if we need to
    assessment_comment_og = assessment_comment,
  ) %>%
  select(-time) %>%
  
  # In the future we should have all of these columns already in the form as per https://github.com/NewGraphEnvironment/dff-2022/issues/119#issuecomment-1781242709
  dplyr::mutate(
    # we need more MoTi sturcture columns so that we can add multiple structure IDs in Q
    moti_chris_culvert_id2 = NA_integer_,
    moti_chris_culvert_id3 = NA_integer_,
    # we may as well have a place for priority follow up ranking and citation keys too.
    # Priority ranking are no fix, low, medium and high, then justify why in the assessment comments
    my_priority = NA_character_,
    my_citation_key1 = NA_character_,
    my_citation_key2 = NA_character_,
    my_citation_key3 = NA_character_) %>%

  # we want these new columns to land at a logical place in the table so we will reorder them
  dplyr::select(
    site_id,
    crew_members,
    date_time_start,
    my_priority,
    assessment_comment,
    contains("_notes"),
    contains("moti_chris_culvert_id"),
    contains("my_citation_key"),
    everything()) %>%
  arrange(crew_members, date_time_start)
```

## Backup and burn back to geopackage

Now we burn the cleaned copy to QGIS project as new file in the `/data_field/2023/form_pscis_2023.gpkg' format. Finally, we backup the cleaned data to a csv and Rdata file and commit for version control using git.

```{r backup}
# burn the cleaned copy to QGIS project as new file
form_pscis_cleaned %>%
  sf::st_write(paste0(path, '/data_field/2023/form_pscis_2023.gpkg'), append=F, delete_dsn=T)

#backup the cleaned data to a csv and Rdata and commit for version control using git
fpr_sp_gpkg_backup(
  path_gpkg = paste0(path, "/data_field/2023/form_pscis_2023.gpkg"),
  dir_backup = "data/backup/simpcw/",
  update_utm = TRUE,
  update_site_id = TRUE,
  write_back_to_path = FALSE,
  write_to_csv = TRUE,
  write_to_rdata = TRUE,
  return_object = FALSE)
```

