---
title: "0130_pscis_wrangle"
date: "Created: 2025-02-13 | Updated: `r format(Sys.Date(), '%Y-%m-%d')`"
output: 
  html_document:
    code_folding: "hide"
params:
  repo_owner: "NewGraphEnvironment"
  repo_name: "fish_passage_fraser_2023_reporting"
  gis_name: "sern_fraser_2024"
  job_name: "2024-074-sern-fraser-fish-passage"
editor_options: 
  chunk_output_type: console
---

```{r setup, echo=TRUE, include = TRUE}
knitr::opts_chunk$set(echo=TRUE, message=FALSE, warning=FALSE, dpi=60, out.width = "100%", eval = FALSE)
options(scipen=999)
options(knitr.kable.NA = '--') #'--'
options(knitr.kable.NAN = '--')
```


# Purpose of this script

Add columns to `form_pscis_2024` and then burn back a NEW geopackage.

This script is still a WIP and there is an issue about what still needs to be added here https://github.com/NewGraphEnvironment/fish_passage_template_reporting/issues/56


```{r params}
# path to raw form_fiss_site after QA
path_form_pscis_raw <- fs::path_expand(fs::path("~/Projects/gis/", params$gis_name, "/data_field/2024/form_pscis_2024_raw.gpkg"))

#NEW geopcackage path for the form after we have added columns as per this script
path_form_pscis <- fs::path_expand(fs::path("~/Projects/gis/", params$gis_name, "/data_field/2024/form_pscis_2024.gpkg"))
```



## Backup the form after the QA

Name the project directory we are pulling from and import the cleaned form from Q after review and finalization.

We use `fpr_sp_gpkg_backup` to import and backup the data, as well as update the UTMS. We will also write the data to a csv and Rdata file and commit for version control using git. Commit these backup files with a message such as "backups after QA - 2024 data".


```{r import}

# read in cleaned form from Q after review and finalization
# backup to csv and rdata
form_pscis_raw <- fpr::fpr_sp_gpkg_backup(
  path_gpkg = path_form_pscis_raw,
  update_utm = TRUE,
  update_site_id = FALSE, ## Turn this off after adding pscis ids
  write_back_to_path = FALSE,
  write_to_csv = TRUE,
  write_to_rdata = TRUE,
  return_object = TRUE)
```


## Fix times

This is not always applicable and we will hopefully have a permanent fix soon.

For Fraser 2024 - times are even more funky than usual- see issue here https://github.com/NewGraphEnvironment/fish_passage_fraser_2023_reporting/issues/123

The times in `form_pscis_2024_raw` are incorrect in the form they are in UTC - but for some reason when read into R the timezone is set to America/Vancouver even though the times are in UTC.

In addition - the 4 sites I added by hand (issue here https://github.com/NewGraphEnvironment/fish_passage_fraser_2023_reporting/issues/121) are the correct time and timezone - so they don't need any fixing. 



```{r time-fix, eval F}

form_pscis_time_fix <-  form_pscis_raw |> 
 dplyr::mutate(date_time_start_raw = date_time_start) |> 
 dplyr::mutate(date_time_start = dplyr::case_when(time_correct == FALSE ~ lubridate::force_tz(date_time_start_raw, tzone = "UTC"), TRUE ~ date_time_start_raw),
               date_time_start = dplyr::case_when(time_correct == FALSE ~ lubridate::with_tz(date_time_start, tzone = "America/Vancouver"), TRUE ~ date_time_start_raw)) |> 
  dplyr::relocate(date_time_start_raw, .before = date_time_start)


form_pscis_raw <- form_pscis_time_fix |> 
  dplyr::select(-date_time_start_raw)

```


## Clean and prep

Do some more cleaning
- fix some vocabulary
- add in the pscis assessment phase

Crossing 22202142 is a phase 1 and phase 2 site. It needs to be submitted in the pscis_phase1 spreadsheet, then resubmitted in the pscis_phase2 spreadsheet once we have its PSCIS ID.

```{r pscis-clean}

form_pscis_clean <- form_pscis_raw |>
  dplyr::mutate(date_time_start = lubridate::ymd_hms(date_time_start, tz = "America/Vancouver"),
                date = lubridate::date(date_time_start)) |>
  
  # add in pscis phase
  ## DOUBLE CHECK IF THERE ARE ANY SITES THAT ARE PHASE 1 AND 2 SITES!!! FIRST TIME HAB CON SITES
  # Crossing 22202142 is a phase 1 and phase 2 site. It needs to be submitted in the pscis_phase1 spreadsheet, then resubmitted in the pscis phase 2 spreadsheet once we have its PSCIS ID. 
  dplyr::mutate(across(tidyselect::starts_with("assess_"), stringr::str_to_title)) |>

  dplyr::mutate(pscis_phase = dplyr::case_when(
    assess_type_phase1 == "Yes" ~ "phase 1",
    assess_type_phase2 == "Yes" ~ "phase 2",
    assess_type_reassessment == "Yes" ~ "reassessment"),
    
    # Fix the vocabulary
    stream_name = stringr::str_replace_all(stream_name, 'Trib ', 'Tributary '),
    stream_name = stringr::str_to_title(stream_name),
    road_name = stringr::str_to_title(road_name),
    road_name = stringr::str_replace_all(road_name, 'Hwy', 'Highway '),
    road_name = stringr::str_replace_all(road_name, 'Fsr', 'FSR'),
    road_name = stringr::str_replace_all(road_name, 'Rd', 'Road '),
    crew_members = stringr::str_to_upper(crew_members))

```

## Burn back to the geopackage
Burn back to geopackage so all the new columns/data are in the form

```{r fiss-burn-gpkg}
form_pscis_clean |> 
  sf::st_write(dsn = path_form_pscis,
               append = FALSE,
               delete_dsn = TRUE)

```


## Backup the new complete form 

```{r backup-complete}

fpr::fpr_sp_gpkg_backup(
  path_gpkg = path_form_pscis,
  update_utm = FALSE,
  update_site_id = FALSE, ## Turn this off after adding pscis ids
  write_back_to_path = FALSE,
  write_to_csv = TRUE,
  write_to_rdata = TRUE,
  return_object = FALSE)
```




## Add PSCIS Ids

Add the PSCIS Ids once they are in the system and burn back to geopackage.

```{r add-pscis-ids, eval = F}

# Add pscis Ids
form_pscis_ids <- dplyr::left_join(form_pscis_raw,
                         xref_pscis_my_crossing_modelled,
  by = c('my_crossing_reference' = 'external_crossing_reference')) |>
  
  dplyr::mutate(pscis_crossing_id = dplyr::case_when(
    is.na(pscis_crossing_id) ~ as.numeric(stream_crossing_id),
    TRUE ~ pscis_crossing_id
  )) |> 
  dplyr::select(-stream_crossing_id)


# Burn to geopackage
form_pscis_ids |> 
  sf::st_write(dsn = path_form_pscis,
               append = FALSE,
               delete_dsn = TRUE)


#Then backup
fpr::fpr_sp_gpkg_backup(
  path_gpkg = path_form_pscis,
  update_utm = FALSE,
  update_site_id = FALSE, ## Turn this off after adding pscis ids
  write_back_to_path = FALSE,
  write_to_csv = TRUE,
  write_to_rdata = TRUE,
  return_object = FALSE)
```

