---
title: "0150_pscis_export_submission"
date: "Created: 2025-02-25 | Updated: `r format(Sys.Date(), '%Y-%m-%d')`"
output: 
  html_document:
    code_folding: "hide"
params:
  repo_url: 'https://github.com/NewGraphEnvironment/fish_passage_fraser_2023_reporting'
  report_url: 'https://www.newgraphenvironment.com/fish_passage_fraser_2023_reporting'
  repo_name: "fish_passage_fraser_2023_reporting"
  gis_name: "sern_fraser_2024"
  job_name: "2024-074-sern-fraser-fish-passage"
  name_pdf: "fish_passage_fraser_2023_reporting.pdf"
  project_area: "fraser"
  project_year: "2024"
editor_options: 
  chunk_output_type: console
---

```{r setup, echo=TRUE, include = TRUE}
knitr::opts_chunk$set(echo=TRUE, message=FALSE, warning=FALSE, dpi=60, out.width = "100%", eval = FALSE)
options(scipen=999)
options(knitr.kable.NA = '--') #'--'
options(knitr.kable.NAN = '--')
```


# Purpose of this script

Scripts to transfer photos and excel file to the PSCIS submission folder

We also QA to be sure that we have all 6 required photos for submission to PSCIS


# PSCIS Submissions

Functions needed for this script

```{r functions}

tfpr_filter_list <- function(idx){
  filestocopy_list[idx]
}

tfpr_photo_change_name <- function(filenames_to_change = filestocopy_list){
  gsub(filenames_to_change, pattern = path_photos, replacement = targetdir)
}
```


## Phase 1

### Move Photos

Move photos to Onedrive

```{r copy-photos}

name_submission <- 'pscis_phase1.xlsm'
name_folder <- paste0("PSCIS_", params$project_area,"_", params$project_year, "_phase1")

# Create folders and copy over photos -------------

targetdir = fs::path_expand(fs::path("~/Library/CloudStorage/OneDrive-Personal/Projects/submissions/PSCIS/", params$project_year, params$project_area, name_folder))

# create the directory. Take note about which phase
fs::dir_create(targetdir)


# use the pscis spreadsheet to make the folders to copy the photos to. For Fraser 2024, the photos are stored in Onedrive.

path_photos = fs::path_expand(fs::path("~/Library/CloudStorage/OneDrive-Personal/Projects/", params$job_name, "data/photos"))

d <- fpr::fpr_import_pscis(workbook_name = name_submission)


folderstocopy<- d$my_crossing_reference |> as.character()

path_to_photos <- fs::path(path_photos, folderstocopy)

# here we transfer just the photos with labels over into the PSCIS directory where we will upload from to the gov interface

folderstocreate<- fs::path(targetdir, folderstocopy)

##create the folders
fs::dir_create(folderstocreate)


# Identify photos that should be copied over into file
filestocopy_list <- path_to_photos |>
  purrr::map(fpr::fpr_photo_paths_to_copy) |>
  purrr::set_names(basename(folderstocreate))


##view which files do not have any photos to paste by reviewing the empty_files object
empty_idx <- which(!lengths(filestocopy_list))
empty_files <- empty_idx |> tfpr_filter_list()

##rename long names if necessary

photo_sort_tracking <- path_to_photos |>
  purrr::map(fpr::fpr_photo_document_all) |>
  purrr::set_names(folderstocopy) |>
  bind_rows(.id = 'folder') |>
  mutate(photo_name = str_squish(str_extract(value, "[^/]*$")),
         photo_name_length = stringr::str_length(photo_name))

##here we back up a csv that gives us the new location and name of the original JPG photos.

##burn to csv
photo_sort_tracking |>
  readr::write_csv(file = "data/backup/photo_sort_tracking_phase1.csv")

## change path name so we can paste to folders
filestopaste_list <- filestocopy_list |>
  map(tfpr_photo_change_name)

##!!!!!!!!!!!!!!!copy over the photos!!!!!!!!!!!!!!!!!!!!!!!
mapply(fs::file_copy,
       path =  filestocopy_list,
       new_path = filestopaste_list)

```


### QA Photos

QA to be sure that we have all 6 required photos for submission to PSCIS.

```{r photo-qa}

# do a little QA to be sure all the photos are there.

t <- fpr::fpr_photo_qa_df(dat = d, dir_photos = fs::path(targetdir, "/"))

# t shows any MISSING photos - no missing photos in this case

```


### Move Files

Move Pscis file

```{r move-pscis-file}

fs::file_copy(path = fs::path('data', name_submission),
              new_path = fs::path(targetdir, name_submission),
              overwrite = T)
```


Make a little readme for the pdf for upload to ecocat and other details

```{r make-readme}

#make a little readme for the pdf for upload to ecocat and other details
writeLines(
  paste(
    "Online interactive report is located at: ",
    params$report_url,
    "",
    "A versioned PDF of the report can be downloaded from: ",
    paste0(params$repo_url, "/raw/main/docs/", params$name_pdf),
    "",
    "Raw data is available here: ",
    paste0(params$repo_url, "/blob/main/data"),
    "",
    "All scripts to produce the online interactive report and PDF are located at: ",
    params$repo_url,
    sep = "\n"
  ),
  fs::path(targetdir, "readme.txt")
)


```



## Phase 2

### Move Photos

Move photos to Onedrive

```{r copy-photos}

name_submission <- 'pscis_phase2.xlsm'
name_folder <- paste0("PSCIS_", params$project_area,"_", params$project_year, "_phase2")

# Create folders and copy over photos -------------

targetdir = fs::path_expand(fs::path("~/Library/CloudStorage/OneDrive-Personal/Projects/submissions/PSCIS/", params$project_year, params$project_area, name_folder))

# create the directory. Take note about which phase
fs::dir_create(targetdir)


# use the pscis spreadsheet to make the folders to copy the photos to. For Fraser 2024, the photos are stored in Onedrive.

path_photos = fs::path_expand(fs::path("~/Library/CloudStorage/OneDrive-Personal/Projects/", params$job_name, "data/photos"))

d <- fpr::fpr_import_pscis(workbook_name = name_submission)


folderstocopy<- d$my_crossing_reference |> as.character()

path_to_photos <- fs::path(path_photos, folderstocopy)

# here we transfer just the photos with labels over into the PSCIS directory where we will upload from to the gov interface

folderstocreate<- fs::path(targetdir, folderstocopy)

##create the folders
fs::dir_create(folderstocreate)


# Identify photos that should be copied over into file
filestocopy_list <- path_to_photos |>
  purrr::map(fpr::fpr_photo_paths_to_copy) |>
  purrr::set_names(basename(folderstocreate))


##view which files do not have any photos to paste by reviewing the empty_files object
empty_idx <- which(!lengths(filestocopy_list))
empty_files <- empty_idx |> tfpr_filter_list()

##rename long names if necessary

photo_sort_tracking <- path_to_photos |>
  purrr::map(fpr::fpr_photo_document_all) |>
  purrr::set_names(folderstocopy) |>
  bind_rows(.id = 'folder') |>
  mutate(photo_name = str_squish(str_extract(value, "[^/]*$")),
         photo_name_length = stringr::str_length(photo_name))

##here we back up a csv that gives us the new location and name of the original JPG photos.

##burn to csv
photo_sort_tracking |>
  readr::write_csv(file = "data/backup/photo_sort_tracking_phase2.csv")

## change path name so we can paste to folders
filestopaste_list <- filestocopy_list |>
  map(tfpr_photo_change_name)

##!!!!!!!!!!!!!!!copy over the photos!!!!!!!!!!!!!!!!!!!!!!!
mapply(fs::file_copy,
       path =  filestocopy_list,
       new_path = filestopaste_list)

```


### QA Photos

QA to be sure that we have all 6 required photos for submission to PSCIS.

```{r photo-qa}

# do a little QA to be sure all the photos are there.

t <- fpr::fpr_photo_qa_df(dat = d, dir_photos = fs::path(targetdir, "/"))

# t shows any MISSING photos - no missing photos in this case

```


### Move Files

Move Pscis file

```{r move-pscis-file}

fs::file_copy(path = fs::path('data', name_submission),
              new_path = fs::path(targetdir, name_submission),
              overwrite = T)
```


Make a little readme for the pdf for upload to ecocat and other details

```{r make-readme}

#make a little readme for the pdf for upload to ecocat and other details
writeLines(
  paste(
    "Online interactive report is located at: ",
    params$report_url,
    "",
    "A versioned PDF of the report can be downloaded from: ",
    paste0(params$repo_url, "/raw/main/docs/", params$name_pdf),
    "",
    "Raw data is available here: ",
    paste0(params$repo_url, "/blob/main/data"),
    "",
    "All scripts to produce the online interactive report and PDF are located at: ",
    params$repo_url,
    sep = "\n"
  ),
  fs::path(targetdir, "readme.txt")
)


```




## Reassessments

### Move Photos

Move photos to Onedrive

```{r copy-photos}

name_submission <- 'pscis_reassessments.xlsm'
name_folder <- paste0("PSCIS_", params$project_area,"_", params$project_year, "_reassessments")

# Create folders and copy over photos -------------

targetdir = fs::path_expand(fs::path("~/Library/CloudStorage/OneDrive-Personal/Projects/submissions/PSCIS/", params$project_year, params$project_area, name_folder))

# create the directory. Take note about which phase
fs::dir_create(targetdir)


# use the pscis spreadsheet to make the folders to copy the photos to. For Fraser 2024, the photos are stored in Onedrive.

path_photos = fs::path_expand(fs::path("~/Library/CloudStorage/OneDrive-Personal/Projects/", params$job_name, "data/photos"))

d <- fpr::fpr_import_pscis(workbook_name = name_submission)


folderstocopy<- d$my_crossing_reference |> as.character()

path_to_photos <- fs::path(path_photos, folderstocopy)

# here we transfer just the photos with labels over into the PSCIS directory where we will upload from to the gov interface

folderstocreate<- fs::path(targetdir, folderstocopy)

##create the folders
fs::dir_create(folderstocreate)


# Identify photos that should be copied over into file
filestocopy_list <- path_to_photos |>
  purrr::map(fpr::fpr_photo_paths_to_copy) |>
  purrr::set_names(basename(folderstocreate))


##view which files do not have any photos to paste by reviewing the empty_files object
empty_idx <- which(!lengths(filestocopy_list))
empty_files <- empty_idx |> tfpr_filter_list()

##rename long names if necessary

photo_sort_tracking <- path_to_photos |>
  purrr::map(fpr::fpr_photo_document_all) |>
  purrr::set_names(folderstocopy) |>
  bind_rows(.id = 'folder') |>
  mutate(photo_name = str_squish(str_extract(value, "[^/]*$")),
         photo_name_length = stringr::str_length(photo_name))

##here we back up a csv that gives us the new location and name of the original JPG photos.

##burn to csv
photo_sort_tracking |>
  readr::write_csv(file = "data/backup/photo_sort_tracking_reassessments.csv")

## change path name so we can paste to folders
filestopaste_list <- filestocopy_list |>
  map(tfpr_photo_change_name)

##!!!!!!!!!!!!!!!copy over the photos!!!!!!!!!!!!!!!!!!!!!!!
mapply(fs::file_copy,
       path =  filestocopy_list,
       new_path = filestopaste_list)

```


### QA Photos

QA to be sure that we have all 6 required photos for submission to PSCIS.

```{r photo-qa}

# do a little QA to be sure all the photos are there.

t <- fpr::fpr_photo_qa_df(dat = d, dir_photos = fs::path(targetdir, "/"))

# t shows any MISSING photos - no missing photos in this case

```


### Move Files

Move Pscis file

```{r move-pscis-file}

fs::file_copy(path = fs::path('data', name_submission),
              new_path = fs::path(targetdir, name_submission),
              overwrite = T)
```


Make a little readme for the pdf for upload to ecocat and other details

```{r make-readme}

#make a little readme for the pdf for upload to ecocat and other details
writeLines(
  paste(
    "Online interactive report is located at: ",
    params$report_url,
    "",
    "A versioned PDF of the report can be downloaded from: ",
    paste0(params$repo_url, "/raw/main/docs/", params$name_pdf),
    "",
    "Raw data is available here: ",
    paste0(params$repo_url, "/blob/main/data"),
    "",
    "All scripts to produce the online interactive report and PDF are located at: ",
    params$repo_url,
    sep = "\n"
  ),
  fs::path(targetdir, "readme.txt")
)


```



